\documentclass[english,10pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{bbm}
\usepackage{babel}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\newtheorem{definition}{Definition}

\title{Notes on Homodyne Measurement}
\author{Arnab Ghorui}

% shorthands
\newcommand{\rr}{\hat {\textbf{r}}}
\newcommand{\xx}{\hat{\textbf{x}}}
\newcommand{\dd}[1]{\hat{D}_{#1}}
\newcommand{\ham}{\hat H}
\newcommand{\half}{\frac{1}{2}}
\newcommand{\tr}[1]{\text{Tr}\left[{#1}\right]}
\begin{document}
	\maketitle
	
	\section{Notations}
	\begin{itemize}
		\item $\xx= (\hat x_1, \hat p_1,\cdots,\hat x_n, \hat p_n)^T$, vector of cannonical operators.
		\item $\Omega = \bigoplus_{j=1}^n \Omega_1$, where $\Omega_1 = \begin{pmatrix}
			0 & 1 \\ -1 & 0
		\end{pmatrix}$.\\
		Note that, for $n=1$, $[\hat x_i,\hat x_j] = i[\Omega_1]_{ij}$. Compactly, 
		\begin{equation}
			[\xx, \xx^T] = i\Omega, \tag{Canonical Commutation Relation}
		\end{equation}
		 where, think the commutation relation as element wise commutator. Some properties of $\Omega$:
		 \begin{itemize}
		 	\item[*] $\Omega^T=-\Omega$(Anti-symmetric)
		 	\item[*] $\Omega^2 = \mathbbm{1}_{2n} \implies \Omega^T\Omega=\mathbbm{1}_{2n}$ (Orthogonal)
		 \end{itemize}
		\item Borrowing from the optical and field-theoretical terminologies, canonical degrees	of freedom are also referred to as ‘\textit{modes}’.
		\item $\hat a_j = \frac{\hat x_j + \hat p_j}{\sqrt 2}$, annihilation operator.
		\item \textbf{BCH formula: }$e^{A+B}=e^A e^B e^{-\half[A,B]}$	for operators $A,B$ if $[A,[A,B]]=[B,[B,A]]=0$
	\end{itemize}
	
	\section{Prerequisits}
	\subsection{Displacement operators}
	\label{sec:displacement_operators}
	\begin{definition}[Weyl operators]
		\begin{equation}
			\dd{\xi} =e^{i \xi^T \Omega \xx} = e^{i(\hat{x}_1\xi_2-\hat{p}_1\xi_2)}\otimes\cdots\otimes e^{i(\hat{x}_n\xi_{2n}-\hat{p}_n\xi_{2n-1})},
		\end{equation}
		where, $\xi \in \mathbbm{R}^{2n}$.
	\end{definition}
	
	\textbf{Properties:}	\begin{itemize}
		\item $\dd{\xi}^\dagger \dd{\xi} = \mathbbm{1}$ (Unitary operator).
		\item $\dd{\xi}\dd{\xi}=\dd{2\xi}$.
		\item $\dd{\xi}\dd{\eta} = e^{-\frac{i}{2}\xi^T\Omega \eta}\ \dd{\xi+\eta}$. (\textbf{Prove!})
		\item$\dd{-\bar{\xi}} \xx \dd{\bar{\xi}} = \xx - \bar{\xi}$\\
		\textit{Proof:} For the $k^\text{th}$ component of $\xx$ \textit{i.e.} $\hat x_k$, we have, \begin{align*}
			\dd{-\xi} \hat x_k \dd{\xi} &= e^{-i\xi^T \Omega\xx}\hat x_k e^{i\xi^T \Omega\xx} \\
			&= \hat x_k -i\left[\xi^T\Omega\xx, \hat x_k \right] + \frac{i^2}{2}\left[ \xi^T\Omega\xx, \left[\xi^T\Omega\xx, \hat x_k \right]   \right] + \cdots \hfil (\textit{using BCH})\\
			&=\hat x_k -\xi_k
		\end{align*}
		From here the result follows directly.
		\item $\dd{-\bar{\xi}} = \dd{\bar{\xi}}^\dagger$.
	\end{itemize}
	
	
	
	\subsection{Symplectic Group}
	\textbf{TODO:} Linear canonical transformation and Symplectic group, Canonical transformations are those which respect  \textbf{CCR}.
	\begin{definition}
		[Symplectic group]
		\begin{equation}
			\label{eq:symplectic_group}
			S\in Sp_{2n,\mathbbm{R}} \iff S\Omega S^T=\Omega
		\end{equation}
	\end{definition}
	\textbf{Evolution:}
	
	
	\subsection{Normal Modes}
	\textbf{TODO: }Definition, etc.
	
	\section{Gaussian States}
	
	\subsection{Quadratic Hamiltonian and evolution}
	
	The most general quadratic/second-order hamiltonian can be written as follows.
	\begin{equation}
		\label{eq:quad_ham}
		\ham = \half \xx^T H \xx + \xx^T \mathbf{\xi}.
	\end{equation}
	Here, $\mathbf{\xi}$ is a $2n$-dimensional real vector. $H$ is a $2n\times 2n$ symmetric matrix called \textit{Hamiltonian matrix}, not to be confused with Hamiltonian. It can always be taken as  a symmetric matrix because, the antisymmetric part with give a term proportional to identity matrix due to \textbf{CCR}, which can always be discarded. If we take $\bar{\mathbf{\xi}}=H^{-1}\xi$, then $\hat{H}' = \half (\xx-\bar{\xi})^T H (\xx-\bar{\xi})$ is equivalent to $\hat{H}$ up to some additive constant term. Using the fourth property from section \ref{sec:displacement_operators} we can write,
	\begin{align}
		\hat{H}' = \half (\xx-\bar{\xi})^T H (\xx-\bar{\xi}) &= \half \sum_{jk} (\xx-\xi) _j H_{jk} (\xx-\xi) _j \notag \\
		&= \half \sum_{jk} (\dd{-\xi} \hat x_j \dd{\xi})  H_{jk} (\dd{-\xi} \hat x_k \dd{\xi})  \notag\\
		&= \half \sum_{jk} (\dd{-\xi} \hat x_j H_{jk} \hat x_k \dd{\xi}) \notag\\
		\hat{H}'&= \half \dd{-\bar{\xi}} (\xx^T H \xx) \dd{\bar{\xi}}
	\end{align} 
	One takeaway from  the above is that, it's enough to study the property of the Hamiltonian, $\hat{H} = \half \xx^T H \xx$.
	
	%\vspace*{-\abovedisplayskip}
	\textbf{Evolution of quadratures under free Hamiltonian:} 
	\begin{itemize}
		\item Heisenberg picture of evolution:
		\label{eq:heisenberg}
		\begin{equation}
			 \dot{\hat O} = i[\hat{H}, \hat{O}].
		\end{equation}
		\item Using the above we get, time evolution of quadratures as, $\dot{\xx} = \Omega H \xx$. Solving which one gets,
		\begin{equation}
			\label{eq:free_evolution}
			 \xx(t) = e^{\Omega H} \xx(0)
		\end{equation}
		Since it is an unitary the \textbf{CCR} must be preserved.
		\begin{align}
			i\Omega = [\xx(0), \xx(0)^T] = [\xx(t), \xx(t)^T]  =[e^{\Omega H t} \xx(0), \left(e^{\Omega H t} \xx(0)\right)^T] \\
			= e^{\Omega H t}[\xx, \xx^T] \left(e^{\Omega H t} \right)^T = i e^{\Omega H t} \Omega  \left(e^{\Omega H t} \right)^T
		\end{align}
		So, we see, that for if \textbf{CCR} is to be preserved,  we must have, $e^{\Omega H} \in Sp_{2n,\mathbbm R}$. See eq. [\ref{eq:symplectic_group}].
		\item We can also write , $\xx(t) = e^{i\hat{H} t} \xx(0) e^{-i\hat{H} t}$, because unitary evolution. Denote, $\hat{S}_H = e^{i\hat{H}} = e^{\frac{i}{2} \xx^T H \xx}$ and $S=e^{\Omega H}\in Sp_{2n, \mathbbm{R}}$. Combining the above two result we write,
		\begin{equation}
			\label{eq:symplectic_evolution}
			\hat{S}_H \xx \hat{S}_H^\dagger = S\xx
		\end{equation}
		
	\end{itemize}
	
	\subsection{Gaussian state}
	\begin{definition}[Gaussian State]
		Gaussian states are defined as all the ground and thermal states of second-order Hamiltonians [eq.\ref{eq:quad_ham}] with positive definite Hamiltonian matrix H > 0.
	\end{definition}
	Thus a \textit{Gaussian state} can be written as,
	\begin{equation}
		\label{eq:gaussian_state}
		\rho_G = \frac{e^{-\beta \hat{H}}}{\tr{e^{-\beta \hat{H}}}},
	\end{equation}
	where, $\beta >0$ and $\hat{H}$ is defined in Eq. \ref{eq:quad_ham}. Ground state is the limiting value,
	\begin{equation}
		\label{eq:ground_state}
		\rho_G = \lim_{\beta\to \infty} \frac{e^{-\beta \hat{H}}}{\tr{e^{-\beta \hat{H}}}}.
	\end{equation}
	\textbf{Note:}
	\begin{itemize}
		\item All Gaussian states are mixed state by construction, except for the ground state.
		\item Gaussian states are parametrized by $\beta,\ \mathbf{\xi}$ and $H$. Though $\beta$ is redundant and can be absorbed into $H$, it allows one to single out pure Gausian states as a limiting case like in Eq. \ref{eq:ground_state}. 
		\item Gaussian states can be generated First and second moment of quadrature. We'll talk about them later.
	\end{itemize}
	
	\section{Gaussian operations}
	Gaussian operations are CP-maps those take Gaussian states to Gaussian states.
	\subsection{Gaussian Unitaries}
	One may write most general second order hamiltonian of $n$-modes as, $\hat{H}=\half \dd{-\bar{\xi}} (\xx^T H \xx )\dd{\bar{\xi}} $.
	First note that, $(\dd{-\bar{\xi}} \xx \dd{\bar{\xi}}) (\dd{-\bar{\xi}} \xx \dd{\bar{\xi}}) = \dd{-\bar{\xi}} \xx^2 \dd{\bar{\xi}}$. Then, 
	it is clear that, $ e^{i\hat{H}} =  \dd{-\bar{\xi}} e^{\frac{i}{2}\xx^T H \xx } \dd{\bar{\xi}}$.
	
	%\vspace{.1cm}\\
	Let's introduce some notations and relations:
	\begin{itemize}
		\item  $\hat S = e^{\frac{i}{2} \xx^T H \xx}$\hfill (Free Hamiltonian unitary)
		\item  $S = e^{\Omega H} \in Sp_{2n,\mathbbm{R}}$\hfill(Evolution of quadratures, Eq. \ref{eq:free_evolution})
		\item $\hat S \xx \hat S^\dagger = S \xx$ \hfill (Eq. \ref{eq:symplectic_evolution})
	\end{itemize}
			Since $\hat S$ is unitary,
	\begin{align}
		\hat S (i\xi^T \Omega \xx)^k \hat S^\dagger &= \hat S (i\xi^T \Omega \xx) \hat S^\dagger \hat S (i\xi^T \Omega \xx) \hat S^\dagger \cdots \hat S ^ \dagger\hat S (i\xi^T \Omega \xx) \hat S^\dagger
	\end{align}
	Now each term can be simplified as,
	\begin{align}
		\hat S (i\xi^T \Omega \xx)\hat S ^\dagger &= i \sum_{jk} \hat{S}(\xi_j \Omega_{jk}\xx_k)\hat S ^\dagger \\
		&= i\sum_{jk}\xi_j \Omega_{jk}(\hat{S}\xx_k \hat{S}^\dagger)\\
		&= i \xi^T\Omega(\hat S \xx \hat S^\dagger)\\
		&= i \xi^T\Omega  S \xx
	\end{align}
	Thus we can write, $ \hat S e^{i\xi^T \Omega \xx} \hat S ^\dagger = e^{i\xi^T \Omega S \xx} = e^{i\xi^T S^{-T} \Omega \xx} = \dd{S^{-T}\xi}$.\\
	
	\textbf{TODO: }To show, that any arbitrary quadratic unitary transformation can be generated by Displacement and Symplectic rotations.
	



\end{document}